---
title: "Final Project"
author: "Hannah Hill, Shelby Lauter, Maitreyi Natarajan, Imani Sherrill"
date: "4/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

[Data Source](https://data.ca.gov/dataset/ca-wic-apl-2022-03-03)

rural defined as: Alpine, Amador, Butte, Calaveras, Colusa, Del Norte, El Dorado, Glenn,
Humboldt, Imperial, Inyo, Lake, Lassen, Madera, Mariposa, Mendocino, Merced, Modoc,
Mono, Monterey, Napa, Nevada, Placer, Plumas, San Benito, San Luis Obispo, 
Santa Barbara, Shasta, Sierra, Siskiyou, Solano, Sonoma, Sutter, Tehama, Trinity,
Tulare, Tuolumne, Yolo, Yuba
[Source](https://www.rcrcnet.org/counties)

urban defined as: Alameda, Contra Costa, Fresno, Los Angeles, Orange, Riverside, 
Sacromento, San Bernardino, San Diego, San Francisco, San Joaquin, San Mateo,
Santa Clara, Ventura
[Source](https://urbancounties.com/about/)

# 1. Data Loading and Cleaning
```{r 1, warning=FALSE}

library(tidyverse)
library(lubridate)
library(dplyr)
library(anytime)

#1.a data setup
wic <- read.csv("wic_redemption.csv") %>%
janitor::clean_names() 
wic <- wic %>%
rename(n_families_redeemed = number_of_families_redeemed_food_benefits) %>%
rename(dollar_amount_redeemed = dollar_amount_redeemed_for_food_instruments) %>%
rename(year_month = obligation_year_and_month) %>%
filter(year_month != 2019 & year_month !=2020)

#create date variable
wic_date <- wic %>%
    mutate(date = ym(year_month)) %>%
   mutate(year = year(date))

#versions 1, 2, and 3. Still getting the column as unknown. We don't have a time component, so I am not sure how strftime will work in this. Also it's an unknown class not POSIXlt
wic_date2 <- wic_date %>%
  mutate(date2 = as_date(x=date, format = '%Y-%m-%d'))

wic_date2 <- wic_date %>%
  filter(!is.na(date)) %>%
  mutate(date2 = as.Date(date))

wic_date2 <- wic_date %>%
 as.Date(x=date, format = '%Y-%m-%d')



#fix dollar amounts; use string function to get rid of $ and ,
wic_date3 <- wic_date2 %>%
  as.numeric(dollar_amount_redeemed)


library(tidymodels)

#1.b split the data
set.seed(20201020)
wic_split <- initial_split(data = wic_date, prop = 0.75)
wic_train <- training(x = wic_split)
wic_test <- testing(x = wic_split)

#1.c EDA - use groupby() and count to determine by group
wic_train %>%
  ggplot(aes(y=dollar_amount_redeemed, x=date)) + 
  geom_point() +
  labs(title = "Study of Family Redemption of WIC Benefits in California") +
  xlab("Level") +
  ylab("Number of Families that Redeemed WIC Benefits") +
  theme(axis.text.y = element_text(vjust = 0.5, hjust=1))

#recipe
wic_recipe <- recipe(
  date ~ ., data = wic_train) %>%
  step_integer(all_predictors()) %>%
  step_nzv(all_predictors()) %>%
  step_normalize(all_predictors())
```

# 2. Machine Learning
```{r 2. warning=FALSE}

library(rsample)
library(parsnip)
library(recipes)
library(workflows)
library(tune)
library(yardstick)

#implement one candidate model
tree_mod <-
  rand_forest(mode = "regression", engine = "ranger")

#create a workflow
tree_wf <- workflow() %>%
  add_recipe(chi_rec) %>%
  add_model(tree_mod)

#fit model
tree_fit <- tree_wf %>%
  fit_resamples(resamples = folds, metrics = metric_set(rmse, mae),
                control = control_resamples(save_pred = TRUE))

#find best rmse
collect_metrics(tree_fit, summarize = TRUE)

#plot the rmse
collect_metrics(tree_fit, summarize = FALSE) %>%
filter(.metric == "rmse") %>%
ggplot(aes(id, .estimate, group = .estimator)) +
geom_line() +
geom_point() +
scale_y_continuous() +
labs(title = "Calculated RMSE Across the 10 Folds",
y = "RMSE_hat") +
theme_minimal()



```